A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors

### 2024-11-05 hybrid-nts.tex
Further chapter 5 content

\subsection{1-D Neutronics Numerical Results}

@@

\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}

@@

\begin{table}[htb]
  \centering
  \caption{List of reactor dimensions.}
  \begin{tabular}{l l}
    Height [cm] & 215 \\
    TODO: Populate table & \\
  \label{table:reactor-dimensions}
\end{table}

@@

\subsection{3-D Numerical Results} \label{sec:3d-nts}

@@

The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
withdrawn. Table \ref{table:initial-crit}
shows the $k_\text{eff}$ values from \gls{MSRE} experimental data, numerical benchmark data
\cite{fratoni_molten_2020}, and this work. All numerical models overestimate the $k_\text{eff}$
relative to the experimental value. In the \gls{MSRE} numerical benchmark report, the authors
attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
simulations in this work used the same
ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.

@@

The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
the Serpent 2 model from the numerical benchmark.
Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
particularly small near the control rods due to strong thermal neutron capture by Gadolinium in the
rods.

@@

Due to the experimental procedure of computing integral rod worth from cumulative differential rod
worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
Both OpenMC and the hybrid method report consistently larger integral rod worths than the
\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
integral rod worth.

Table \ref{table:rod-worth} shows the total rod worth of Rod 1.

\begin{figure}[p]
  \centering
  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
  \label{fig:rod-worth}
\end{figure}

\begin{table}[htb]
  \centering
  \caption{Total rod worth of Rod 1 when fully inserted.}
  \begin{tabular}{l S[table-format=4(2)] S}
    \toprule
     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
     \cmidrule(l){2-2}
    \gls{MSRE} experimental data & 2250(67) & {-}\\
    OpenMC (This work) & 2364(44) & 5.1 \\
    Hybrid (This work) & 2345 & 4.2 \\
    Diffusion (This work) & 2725 & 21.1 \\
    \bottomrule
  \end{tabular}
  \label{table:rod-worth}
\end{table}

@@

@@

>>>>

\subsection{1-D Neutronics Numerical Results \& Discussion}

@@

\subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}

@@

\begin{table}[htb]
  \centering
  \caption{List of reactor dimensions.}
  \begin{tabular}{l l}
    Height [cm] & 215 \\
    TODO: Populate table & \\
  \end{tabular}
  \label{table:reactor-dimensions}
\end{table}

@@

\subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}

@@

The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
withdrawn. Table \ref{table:initial-crit} shows the $k_\text{eff}$ values from \gls{MSRE}
experimental data, numerical benchmark data \cite{fratoni_molten_2020}, and this work. All
numerical models overestimate the $k_\text{eff}$ relative to the experimental value by 1-2\%
(1000-2000 pcm). In the \gls{MSRE} numerical benchmark report, the authors
attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
simulations in this work used the same
ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.

@@

The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
the Serpent 2 model from the numerical benchmark.
Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
particularly small near the control rods due to strong thermal neutron capture cross section of
gadolinium.

@@

Due to the experimental procedure of computing integral rod worth through cumulative differential
rod worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
Both OpenMC and the hybrid method report consistently larger integral rod worths than the
\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
integral rod worth.
Table \ref{table:rod-worth} shows the total rod worth of Rod 1. Similar to observations from Figure
\ref{fig:rod-worth}, OpenMC and the hybrid method report total rod worths marginally larger than
the \gls{MSRE} data by 5.1\% and 4.2\%, respectively, while the diffusion method exceeds the
experimental value by 21.1\%.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
  \label{fig:rod-worth}
\end{figure}

\begin{table}[t]
  \centering
  \caption{Total rod worth of Rod 1 when fully inserted.}
  \begin{tabular}{l S[table-format=4(2)] S}
    \toprule
     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
     \cmidrule(rl){2-2} \cmidrule(l){3-3}
    \gls{MSRE} experimental data & 2250(67) & {-}\\
    OpenMC (This work) & 2364(44) & 5.1 \\
    Hybrid (This work) & 2345 & 4.2 \\
    Diffusion (This work) & 2725 & 21.1 \\
    \bottomrule
  \end{tabular}
  \label{table:rod-worth}
\end{table}

The most likely factor for OpenMC and the hybrid method overpredicting the control rod worth is the
modeling of control rods as solid annular cylinders as opposed to 36 separate control elements
strung along a flexible stainless steel cable. Figure \ref{fig:rod-bead} depicts a poison element
and its physical dimensions. Other than the Gd$_2$O$_3$-Al$_2$O$_3$ poison material, each control
element has a top and bottom closure made of INOR-8 alloy. According to the physical dimensions,
the closures make up 15.1\% of the total element height. The 3-D numerical \gls{MSRE} model in this
work omits the outer and inner INOR-8 tubes and the flexible stainless cable, and replaces the top
and bottom closure with poison material. Fully resolving the closures in Moltres would require
excessively fine axial mesh resolution and advanced adaptive meshing features not currently
available. A straightforward corrective measure for future \gls{MSRE} simulations would be to
retain the solid annular cylinder structure and adjust the rod dimensions or the rod material
composition to match the expected worth from \gls{MSRE} data or OpenMC simulations with the beaded
geometry.

\begin{figure}[t]
  \begin{subfigure}[b]{0.33\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{rod-bead-2}
  \end{subfigure}
  \begin{subfigure}[b]{0.65\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{rod-bead}
  \end{subfigure}
  \caption{Cutaway view and dimensions of a control rod poison element.}
  \label{fig:rod-bead}
\end{figure}

Overall, the results show that the hybrid $S_N$-diffusion method accurately reproduces the control
rod worth at various levels of insertion. As observed in the 1-D and 2-D models, the 3-D model
$k_\text{eff}$ values exhibit some discrepancies relative to OpenMC, but the discrepancies do not
affect rod worth estimates because they are systematic errors. This \gls{MSRE} hybrid model may be
improved in future work by characterizing and applying neutron albedo at the vacuum boundaries.

@@

This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
values measured at 872 K and 972 K with 67.86 $^{235}$U loading. This approach is identical to that
of the \gls{MSRE} experimental measurements \cite{prince_zero-power_1968} and the \gls{MSRE}
numerical benchmark calculations \cite{fratoni_molten_2020}. The temperature reactivity coefficient
is calculated by dividing the difference in reactivities by the temperature change (100 K). All
simulations included the effects of thermal expansion on the material density but left all reactor
physical dimensions unchanged.

Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
results fall within one standard deviation of the \gls{MSRE} data.

\begin{table}[t]
  \small
  \centering
  \caption{Isothermal temperature reactivity coefficients with 67.86 kg $^{235}$U loading from
  \gls{MSRE} data, the \gls{MSRE} numerical benchmark \cite{fratoni_molten_2020}, and the OpenMC
  and Moltres models in this work.}
  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
    \toprule
     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
     \cmidrule(rl){2-2} \cmidrule(l){3-3}
    \gls{MSRE} experimental data & 13.41(150) & {-}\\
    Serpent 2 (Numerical benchmark) & 12.51(23) & -6.7 \\
    OpenMC (This work) & 13.85(43) & 3.3 \\
    Hybrid (This work) & {TBD} &  \\
    Diffusion (This work) & {TBD} &  \\
    \bottomrule
  \end{tabular}
  \label{table:temp-coef}
\end{table}

@@

% Strong scaling tests & discussion

% Difficulties with 3-D modeling
% SN void treatment constants, Transfer times
